% ======================================================================
% Historical and Source Companion (RGCS v3, Agent 09)
% Compile: latexmk -xelatex historical_source_companion.tex
% Generated tables from tools/make_v3_artifacts.py.
% ======================================================================
\documentclass[10pt,letterpaper]{article}
\input{../common/rgcs_v3_preamble.tex}
\usepackage[hidelinks,pdftitle={RGCS Historical and Source Companion},pdfauthor={Andrew Green}]{hyperref}
\graphicspath{{figures/}}
\input{tables/generated_values}
\rhead{\small Historical \& Source Companion}

\title{\vspace{-2.0em}\textbf{Historical and Source Companion\\
to RGCS v3 / RSCS 1.0}}
\author{Andrew Green}
\date{2026}

\begin{document}
\maketitle

\begin{abstract}
This companion preserves the source material behind the RGCS programme
--- historical practice, personal logs, theoretical proposals, and recent
peer-reviewed photonics --- in its own words and with its own authors,
and records exactly what the programme did with each piece: what
mathematics was adapted (always re-derived and re-tested), what
interpretation was \emph{excluded} by a binding forbidden-transfer
clause, and which testable hypotheses resulted. Nothing here is evidence
for anything; that is the point of keeping it separate. The register
covers \gvNumSources{} sources and \gvNumEp{} equation-level adaptation
rows.
\end{abstract}

\section{Why a companion}
Programmes that draw inspiration from unconventional sources face two
symmetric failure modes: laundering (source claims quietly become
physics by repetition) and erasure (the historical material is discarded
or ridiculed, destroying the record of where questions came from). The
RGCS discipline avoids both with one rule: \textbf{every statement keeps
its author and its class}. A source claim \SRC{} may inspire a
hypothesis \HYP{} only through an explicit, registered translation with
an observable, controls, and a failure condition --- and the original
wording is preserved here unchanged.

\section{The historical crystal-practice corpus \SRC}
The programme's origin material is a practice tradition around faceted
quartz instruments (the Vogel-style cut lineage) and a personal
working log (``JH log'') whose operating points parameterize the v2/v3
protocols: the \SI{4096}{\hertz} coil carrier; the sound keys at 1496,
644, and \SI{587}{\hertz}; the \SI{20}{\hertz} electrode-pulse family
with its 15\,V setting; the 46\,ms burst envelopes with their spacings
and pauses; opposed complementary coils; and ``the eye'' --- a special
interior locus of the crystal. In the source's own frame these are
statements of practice, made in earnest, with their own vocabulary of
resonance, keys, and awakening. The programme records them respectfully
and verbatim in the source-evidence ledger, assigns each an RG-$\ast$
identifier, and gives them \emph{no truth value}: they select where to
look, never what is true.

Two translations illustrate the method. ``A single crystal tone''
becomes the scalar effective wave speed $\vL$ \eqid{RGCS-M.10} --- which
v3's anisotropic model then resolves, recovering the scalar as the
orientation-unknown case. ``The eye'' becomes an eight-entry node menu
(metric centre, shaft prior, measured vibration node, electrical node,
optical feature, overlap maximum, coupling maximum, phase singularity),
of which the measured node supersedes all others and five entries are
pre-registered hypotheses H-24\,--\,H-28 with failure conditions.

\section{The Arisaka NHT/HAL proposals \SRC}
The Neural Holography Theory and Holographic Ring Attractor Lattice
proposals \citep{arisaka2022nht,arisaka2022vision,arisaka2023colloquium}
are theoretical neuroscience: space encoded in the phase of oriented
traveling waves, ring-attractor lattices as memory, egocentric
$\leftrightarrow$ allocentric transforms. They are their authors'
proposals and remain so. The programme adapted their \emph{structure}
--- a typed memory record with allocentric/egocentric positions, a frame
transform, predicted vs.\ observed states \eqid{RSCS-C.15}, \ENG{} ---
which is useful software regardless of the neuroscience, while the
space-to-phase encoding \eqid{RSCS-O.3} and memory lattice
\eqid{RSCS-C.14} are quarantined \HYP{} requiring an explicit
acknowledgment flag to construct. \textbf{Excluded outright}: any
consciousness claim, any brain-mechanism claim, and any claim that
quartz realizes memory.

\section{The photonics source set \SRC{} $\to$ adapted mathematics}
Six recent works supplied the optical/nonreciprocal mathematics:
phonon-mediated Autler--Townes splitting \citep{sohn2021phonon},
terahertz-bandwidth acousto-optic isolation \citep{cheng2024terahertz},
self-induced nonlinear nonreciprocity \citep{wang2025selfinduced},
noiseless single-photon isolation \citep{zhang2023noiseless}, TMOKE
coupled-waveguide beating \citep{chao2025tmoke}, and latched-garnet
Faraday isolation via spiral-helix laser writing
\citep{lapointe2023faraday}. Each achieves its effect through an active
ingredient quartz is not claimed to have; the binding posture is that an
unbiased passive quartz path is \emph{reciprocal} and every directional
observable's pre-registered expectation is the null. The complete
equation-level ledger --- what crossed, what is forbidden --- follows.
\input{tables/adaptation_ledger}

\section{Evidence boundaries, stated plainly}
As of this writing: no experiment in this programme has confirmed any
non-standard physical hypothesis. The v2 measurement campaigns produced
calibrations, controls, artifact identifications, and negative results
--- all registered. The value the historical material has contributed is
real but methodological: it generated precise, falsifiable questions
that conventional instrumentation can answer. If every pre-registered
hypothesis fails, the companion still stands as an honest record of a
tradition, its language, and a disciplined attempt to test it.

\bibliographystyle{unsrtnat}
\bibliography{../../references/references,../common/handbooks}

\end{document}
